\documentclass{article}

\usepackage[english]{babel} % localisation 

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{csquotes}       % For biblatex with babel
\usepackage[backend=biber,style=authoryear,sorting=nyt, doi=false,isbn=false,url=false,eprint=false]{biblatex} % Package for bibliography (citing)
\bibliography{bibliography.bib}


\begin{document}
\section{Introduction}
\label{sec:intro}

As cities expand, becoming larger and larger,
urban mobility is posing growing challenges to decision makers:
an increase of travel demand interferes indeed with a limited infrastructure,
causing a degeneration of travel conditions.

One of the biggest issues is congestion,
which creates big inefficiencies in the system as a whole:
according to a recent study \parencite{kim2022congestion},
system inefficiencies leading to congestion
cause in California an average delay of four to five minutes per commuter per day,
amounting to an annual economic loss comparable to six billions USD,
equivalent to 1.67\% of the state's GDP.
These figures describe the effect of a systemic inefficiency,
to which a systemic answer needs to be found.

To address this issue, two main strategies can be adopted:
policies can be implemented to affect either the supply side of traffic (that is, the infrastructure)
or the demand side (that is, the behaviour of the users).
On the one hand, implementing policies on the supply side seems an easy solution.
Supply side solutions anyway draw many, different challenges.
Modifying the infrastructure may indeed be infeasible, for different kinds of constraints;
when feasible, it may be costly, or extremely slow.
On the other hand,
one may implement policies which aim to have an impact on the demand side of traffic,
known under the name of Traffic Demand Management (TDM).
TDM policies aim to influence when and how the road users move,
and with this influence to reduce road usage in peak hours and,
consequently, to reduce congestion.
A wide variety of these strategies exist:
examples may include policies such as implementation of fixed or variable tolls,
congestion pricing, High Occupancy Vehicle (HOV) lanes or parking restrictions,
as well as incentives for mode-switching to,
for instance, public transport or carpooling.

These policies are generally more flexible, and easier to implement,
than an infrastructural change.
Despite this, they require a deep understanding of how congestion is created
to effectively tackle it.
Many actors are indeed involved in how congestion develops:
the users, that choose how and when to travel;
governments, which make decisions regarding policy and infrastructure;
the service providers, who are in charge of implementing these policies and react to the users' choices.

Effectively implementing TDM strategies requires understanding how these agents interact and,
in particular, how the users choose when and how to travel.
This project proposal aims to deepen the knowledge of this topic.
The contribution of the project will be twofold:
on the theoretical side, I aim to improve the current models of congestion forming,
mainly dating back to \cite{d0907f84-e14a-3d98-ad20-759f41491d6e}.
The improvement will be targeted to solving existing issues in the model,
as noted in \cite{https://doi.org/10.1111/iere.12692}.
On the applied side, the model still needs to be properly validated with data,
and reliable values for its parameters need to be found via proper calibration of the model.
This calibration can be done by using several data sources,
including modern ones such as drone imaging \parencite{fonod2025songdo}.

The project proposal is structured as follow:
section~\ref{sec:lit} consists of a review of the existing literature,
describing how congestion theory was developed and pointing out some open issues in the field.
Section~\ref{sec:proj_desc} will explain the contributions of this project to the literature,
what I aim to improve and what methods will be used to achieve the aims.

\section{Literature Review}
\label{sec:lit}

Modelling congestion is not a new issue in the literature.
Early works, such as \cite{Jorgensen1947},
relied on static modelling:
the dependency of traffic on the time of the day was neglected,
typically considering different regimes for peak and off-peak congestion.

This approach was challenged by two seminal papers by William Vickrey (\cite{Vickrey1963, Vickrey1969}),
which gave the first example of what are now known as \textit{dynamic} traffic models,
laying the foundation of bottleneck modelling.
In these models, the traffic is assumed to vary on the time of the day.
According to Vickrey's theory,
the road network has some junctions whose capacity is limited,
meaning that there exists a maximal flow of vehicles that can pass through them at any time.
These junctions are called \textit{bottlenecks}, since they restrict the traffic flow as a bottleneck does with a flowing liquid.
Modelling the traffic reduces thus to modelling the interactions between the users and the bottlenecks they face.

Vickrey's work was later extended by \textcite{de1983stochastic},
which look for a stochastic equilibrium solution,
and \textcite{d0907f84-e14a-3d98-ad20-759f41491d6e},
which take into account elastic demand and better formalize the initial model.

These models better characterize how congestion is generated:
each user travelling through a bottleneck is assumed to face a cost,
depending on some user-dependant parameters.
A user thus decides when to commute based on when its cost is minimized.

The model is thus known under the names of \textit{ADL}, from the names of the authors,
or \(\alpha\)-\(\beta\)-\(\gamma\) model, from the parameters in the cost function.

Many works focused in finding an equilibrium solution when a group of users compete for one or more bottlenecks,
by varying either the composition of the users or the topology of the network.
Originally, \textcite{de1983stochastic} found the solution for a group of homogenous users and a single bottleneck.

Their work has been extended, taking into account increasingly complex networks.
Notable works that do this include \textcite{doi:10.1287/trsc.24.3.217, doi:10.1287/trsc.27.2.148},
which study a Y-shaped network,
as well as \textcite{AKAMATSU2015808} for more complex networks.

Similarly to how more complex road networks have been studied,
recent literature has been increasing the complexity of the population of users.

Initial works, such as \textcite{Vickrey1969}, considered the users to be homogeneous.
Notable examples of works relaxing this hypothesis include \textcite{arnott1988schedule, amirgholy2017analytical}.

For a deeper review of bottleneck modelling, please refer to \textcite{LI2020311};
the proposal will now focus more on the literature relevant for the project and,
in particular, on the problems encountered when validating the model with real data.

\textcite{54d203ee-4bf8-3234-9286-56e4c8b7f5bd} does a first attempt in doing so,
by fitting a discrete choice model on survey data.
The values found and reported by it are widely used in the literature,
as in, for instance, \textcite{d0907f84-e14a-3d98-ad20-759f41491d6e}.
\textcite{https://doi.org/10.1111/iere.12692} presents a review of the parameter values used in the literature.

Furthermore, \textcite{https://doi.org/10.1111/iere.12692} highlights a relevant issue in calibrating the model with data and,
as a consequence, in the whole theory:
when studying data about real travel time patterns (that is, how the travel time across a real bottleneck changes through the day),
observed slopes significantly differ from the theoretically predicted ones.

The author of the article suggests a solution, that is,
to consider a big amount of the travellers as \textit{inframarginal},
travellers who are not willing to change their departure time.
While this solution seemingly solves the issue,
it presents several drawbacks, to be addressed in future research.

After this discrepancy with real data is solved,
the model needs to be properly calibrated.
Numerical simulators of road traffic,
such as METROPOLIS and METROPOLIS2 \parencite{de1997metropolis,RePEc:ema:worpap:2024-03},
already converge to a satisfactory approximation of the theoretical equilibria
\parencite{RePEc:ema:worpap:2024-03},
and can efficiently be employed as a bridge between the theory and its application in policy making.
To converge to meaningful solutions,
they need anyway to rely on a proper description of the population.

The next section is devoted to describing how the project could tackle these issues,
and what results it could provide.

\section{Project Description}
\label{sec:proj_desc}

The proposed project is articulated in two main parts:
the first part will deal with improving the \(\alpha\)-\(\beta\)-\(\gamma\) model,
to solve the issues highlighted by \textcite{https://doi.org/10.1111/iere.12692}.

To achieve this aim, different methods can be employed.

First of all, a different distribution of users can be considered.
Considering \textit{inframarginal} users, as in \textcite{https://doi.org/10.1111/iere.12692},
is an instance of this approach:
by allowing travellers to be inframarginal,
the distribution of the parameters \(\beta\), \(\gamma\) is assumed to have a consistent tail.
But no other features of these distributions are specified.
\textcite{yang2026recovering}, assuming the parameters to be normally distributed,
retrieve the mean and variance of the distribution from data.
But normal distributions do not allow them to validate the method with real data,
as inframarginal travellers are not enough in this setting.
Other distributions can be considered,
to evaluate whether simply modifying the users' preferences could solve the observed issue with real data.

Apart from changing the distribution of users,
structural changes to the model can be made,
trading off its simplicity for an improved adherence to real data.

The first modification that can be made is about the coefficients themselves:
considering users facing costs that do not increase linearly on early, late arrivals may produce a more flexible model,
allowing equilibrium solutions similar to the ones observed in reality.

Lastly, uncertainty can be considered in the cost function.
Introducing uncertainty is not trivial, and can be made at several levels:
considering users with distributed preferences introduces uncertainty at the global level,
and is a problem tackled by several articles.
Note anyway that this is different from introducing uncertainty in the cost function of the single user,
and may lead to a different result, with a different minimization process.
This approach can be taken considering risk aversion,
as done in \textcite{doi:10.1177/0361198118792336}, or neglecting it,
in favour of a simpler model to deal with analytically.

The second part of the project concerns the actual validation of the model with data.
This will involve working with datasets and simulators,
including for instance the Performance Measurement System (PeMS) by the California Department of Transportation (CalTrans),
and the simulator METROPOLIS2 \parencite{RePEc:ema:worpap:2024-03}.
The aim of this part will be to actually calibrate the model,
to find how the parameters are distributed in reality,
in accordance with the existing literature.

Several data sources can be employed to validate the model:
Sources used in \textcite{https://doi.org/10.1111/iere.12692} include the PeMS dataset,
as well as data independently collected from Google Maps.
Apart from these, datasets allowing for more complex network topologies can be employed.
These include data obtained from gps tracking of vehicles \parencite{mazare2012trade},
or from several other methods, including even, for instance, drone imaging \parencite{fonod2025songdo}

Note that this second part is dependant from the first one,
and it is intimately connected to it.
Depending on the results of the first part,
calibration with data can become very simple or infeasible.

\newpage
\printbibliography


\end{document}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End:
