\documentclass{article}

\usepackage[english]{babel} % localisation 

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[hidelinks]{hyperref}
\usepackage{csquotes}       % For biblatex with babel
\usepackage[backend=biber,style=authoryear,sorting=nyt, doi=false,isbn=false,url=false,eprint=false]{biblatex} % Package for bibliography (citing)
\bibliography{bibliography.bib}

\begin{document}
\section{Introduction}
\label{sec:intro}

I am writing this Research Statement to express my interest in pursuing a PhD in Economics, Analytics and Decision Sciences at IMT Lucca.

Although applying for an Economics PhD, you can see from my CV that I don’t have any degree in Economics.
I indeed studied Mathematics for my bachelor's, at Sapienza, and Autonomous Systems Engineering for my master’s, at DTU, Denmark.
During my Bachelor, I was mainly interested in theoretical mathematics such as algebra and topology;
during my master's I mainly focused in theoretical and applied computer science, machine learning and some control theory.

During my master's thesis, I spent a semester conducting research at EPFL, in Lausanne,
working and publishing a paper with André de Palma,
a renowned transport economist I had the luck of meeting there.
The research stay, in which I have virtually briefly been a PhD student,
definitely convinced me to remain in academia:
I really enjoyed being independent doing one’s research,
while being able to discuss issues with others as they arose.
I thus decided to pursue a PhD in Economics, but shifting my focus away from Transport Economics,
mainly since I find the methods used in other fields to be more compelling.

This is why I think the PhD program at IMT perfectly suits me:
I would be able to do appropriate, high-level course work,
to properly address the weaknesses of my technical background,
and start doing research in fields I find the most attracting.
Moreover, I think I could really profit from the interdisciplinarity of the school,
as well as the variety of the courses thought:
I find, for instance, extremely interesting the work done in other research groups such as NETWORKS and SYSMA.
Despite this, I want my main focus to be the analysis of social systems.
This means, from my point of view, studying economics;
I thus think developing a solid background in the discipline will be precious for my future research,
and this is why I am applying to the EADS doctoral program.

While pursuing my PhD, I would anyway like to keep working in Transport Economics:
the work done in my master's thesis \parencite{yang2026recovering}
left us with some open questions \parencite[for some context, see][]{https://doi.org/10.1111/iere.12692}
and some potential ways to address them.
I would thus like to keep
working on this field as a side project,
while shifting my main research focus towards topics I find more interesting.
In particular, I would like to work in topics closer to theoretical microeconomics,
and I find interesting the work done by professors such as Giorgio Gnecco and Irene Crimaldi.

This project will thus try to bridge between the research I have done,
that is in Bottleneck Theory, and the research I would like to do,
using game theory as a topic that connects the two fields.

\section{Background}
\label{sec:background}

As cities expand,
urban mobility poses growing challenges to decision makers:
an increase of travel demand interferes indeed with a limited infrastructure,
causing a deterioration of travel conditions.

One of the biggest issues is congestion,
which creates significant inefficiencies in the system as a whole:
according to a recent study \parencite{kim2022congestion},
system inefficiencies leading to congestion
cause an average delay of four to five minutes per commuter per day in California,
amounting to an annual economic loss comparable to six billions USD,
equivalent to 1.67\% of the state's GDP.
These figures describe the effect of a systemic inefficiency,
to which a systemic answer needs to be found.

To address this problem, many complementary approaches can be taken.

One of these approaches consists in considering very simple road networks,
and studying how road users interact with each other,
with the goal of extending the theory to realistic road networks.
This type of works, forming now a field known as \textit{Bottleneck Theory},
was started by two seminal papers by William Vickrey \parencite{Vickrey1963, Vickrey1969}.
Studies at equilibrium were started by
\textcite{de1983stochastic, d0907f84-e14a-3d98-ad20-759f41491d6e},
yielding a model for congestion at equilibrium under elastic demand,
where users can freely decide their departure time.
For a review of work done in the field, refer to \textcite{LI2020311}.

On the other hand, the network itself can be directly studied,
developing the analysis around spatial preferences (i.e. route choice)
rather than departure time choice.
This approach, often grounded in game theory,
has been widely taken in the literature:
a road network is indeed the prototypical application of congestion games \parencite{Rosenthal1973ACO},
and many studies have been done on them,
including for instance the case in which congestion is repeated in several consequent days \parencite{chen1997boundedly},
and agents learn the traffic condition through various methods \parencite{han2006interactive}.

When having this approach, interesting issues arise.
The issue that is the most known is probably the Braess' paradox \parencite[originally published in German in 1968, translated in][]{braess2005paradox}:
this paradox, according to which adding some roads to a traffic network can slow down the overall traffic flowing through it,
became widely known when, in 1990,
the closure of one of the most congested roads in New York led to a decrease of the overall congestion \parencite{kolata_1990}.
This phenomenon suggests a different strategy for describing the road network:
rather than considering agents as the players of the game,
one could consider a cooperative game in which the players are the roads themselves and,
by deciding not to be available to traffic,
they can minimise the total congestion.
This interesting approach is defined by \textcite{HADAS2017120},
and subsequently developed to efficiently compute a measure for the importance of each road in
\textcite{gnecco2021public,passacantando2021braess}.

Interestingly, both the bottleneck and the network-based approach use,
more or less explicitly, concepts of game theory.
Nonetheless, game theory is less extensively used when directly dealing with bottlenecks,
while the elastic demand of traffic is often neglected when dealing with the traffic infrastructure.

The goal of this project proposal will thus be to bridge the two approaches,
by considering users that, in a game-theoretic formulation,
are allowed to change both their route and departure time.

\newpage
\printbibliography

\end{document}

%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End:
